---
title: "Rt"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

## Stochastic model - Gillespie's algorithm
### SEPAIR model 

#### Initial conditions
```{r}
# devtools::install()
devtools::load_all()
library(Kmisc)
# beta is updated by R0
THETA <- list(beta = 0.3, R0 = 1.2, epsilon = 1/2.5, delta = 1/5, 
            gamma = 1/2.5, bp = 1, ba = 1, fa = 0.3, fd = 1, I0 = 1e2, 
            time_dep_Rt = FALSE)
# this changes the default y0 in the Kmisc package
THETA$Y0 <- c(S = 1e7-THETA$I0, E = 0, P = 0, A = 0, 
              I = THETA$I0, R = 0, CE = 0, CI = 10, CR = 0)
THETA$tend <- 200
usethis::use_data(THETA)

# library(rootSolve)
# fun <- function (x) 0.5*sin(0.1*pi*x)+ 1.2 - 0.9
# curve(fun(x), 0, 40)
# abline(h = 0, lty = 3)
# uni <- uniroot(fun, c(30, 35))$root
# points(uni, 0, pch = 16, cex = 2)
```

### Run the ODE version
```{r}
# tstamp <- format(Sys.time(), "%Y%m%dT%H%M%S")
library(Kmisc)
library(ggplot2)
theme_set(theme_bw())
params <- THETA
tend <- params$tend

params$I0 <- 100
params$Y0 <- c(S = 1e7-params$I0, E = 0, P = 0, A = 0, 
              I = params$I0, R = 0, CE = 0, CI = 0, CR = 0)

tstamp <- format(Sys.time(), "%Y%m%d")
times <- seq(0, tend, by = 1) # daily output for 150 days
library(deSolve)
library(tidyverse)

params$time_dep_Rt <- TRUE
ode(y = params$Y0, times = times, func = ode_sepair, parms = params) %>% 
  as.data.frame() -> out 

out$daily_infected <- c(0, diff(out$CE)) 
out$daily_symptom <- c(0, diff(out$CI)) 
out$daily_confirmed <- c(0, diff(out$CR))
out$daily_Rt <- sapply(0:tend, function(x) get_Rt(x))


# saveRDS(out, paste0("outputs/ode_I0_100_", tstamp, ".rds"))

out_long <- out %>% pivot_longer(-time) 
out_long$name <- factor(out_long$name,
                        levels = c("S", "E", "P", "A", "I", "R", 
                                  "CE", "CI", "CR", "daily_infected", 
                                  "daily_symptom", "daily_confirmed", "daily_Rt"))
out_long %>%
  filter(name == "daily_infected" | 
           name == "daily_symptom" | name == "daily_confirmed") %>%
  ggplot(aes(x = time, y = value, color = name)) +
  geom_line(size = 1.2) +
  labs(x='Time (day)', y = 'Number of individuals', color = "") + 
  theme_grey(base_size = 16) 
  # facet_wrap(vars(name), nrow = 2, scales = "free_y")
```

### Run the Gillespie's direct method version
```{r}
# library(Kmisc)
# library(tidyverse)
set.seed(42)
tstart <- Sys.time()
# tstamp <- format(Sys.time(), "%Y%m%d")

nrun <- 2000

# params <- THETA
# params$time_dep_Rt <- TRUE
# tend <- params$tend
# y0 <- params$Y0

res <- gillespie_run(func = gillespie_sepair, 
                     tend = tend, 
                     nrun = nrun, 
                     y0 = params$Y0, 
                     params = params, 
                     report_dt = 1)
Sys.time() - tstart

daily_infected <- data.frame(matrix(0, nrow = tend + 1, ncol = nrun))
daily_symptom <- data.frame(matrix(0, nrow = tend + 1, ncol = nrun))
daily_confirmed <- data.frame(matrix(0, nrow = tend + 1, ncol = nrun))

for (i in seq_len(length(res))) {
  inf <- diff(res[[i]]$CE)
  onset <- diff(res[[i]]$CI)
  conf <- diff(res[[i]]$CR)
  daily_infected[1:(length(inf)+1), i] <- c(0, inf)
  daily_symptom[1:(length(onset)+1), i] <- c(0, onset)
  daily_confirmed[1:(length(conf)+1), i] <- c(0, conf)
}

df <- daily_infected
df$time <- 0:tend
df %>% pivot_longer(cols = -time) -> df
dfrm <- data.frame(time = 0:tend, dinf = rowMeans(daily_infected))

plt <- ggplot() +
  geom_line(data = df, aes(time, value, group = name), 
            color = "grey80") +
  geom_line(data = out, aes(time, daily_infected), # ODE model output
            color = "darkred",
            size = 1, inherit.aes = FALSE) + 
  geom_line(data = dfrm, aes(time, dinf), # ODE model output
            linetype = "dotted",
            size = 1, inherit.aes = FALSE) + 
  labs(x="Time", y="Daily infected cases", color="") + 
  geom_segment(aes(x=c(0,0,0), xend=c(5,5,5), y=c(2000,1800,1600),
                yend=c(2000,1800,1600)), size=c(1,1,1), 
               color = c("grey80", "darkred", "black"),
               linetype = c("solid","solid","dotted")) +
  geom_text(aes(x=c(7,7,7), y=c(2000,1800,1600)), size=c(4,4,4), 
               hjust = c(0,0,0), vjust = c(0.5,0.5,0.5), 
               label = c("Stochastic simulations", "ODE", "Mean of the stochastic simulations"))
# plt            
# ggsave("plots/ODE_stoch_IO_100_infected.png", plt)

stoch <- list()
stoch$daily_confirmed <- daily_confirmed
stoch$daily_symptom <- daily_symptom
stoch$daily_infected <- daily_infected
stoch$daily_Rt <- sapply(0:tend, function(x) get_Rt(x))

# saveRDS(stoch, paste0("outputs/stoch_I0_100_", tstamp, ".rds"))
```

### Apply a detection rate
```{r}
rr <- nrow(stoch$daily_confirmed)
cc <- ncol(stoch$daily_confirmed)
detected <- data.frame(matrix(0, nrow = rr, ncol = cc))
stoch_partial_obs <- list()
probs <- seq(0.2, 0.8, 0.1)
stoch_partial_obs$probs <- probs
set.seed(42)
for (k in 1:length(probs)) {
  prob <- probs[k]
  for (i in 1:rr) {
     for (j in 1:cc){
       size <- stoch$daily_confirmed[i, j]
       # message(paste0("i = ", i, ", j = ", j, ", case = ", size, "\n"))
       if (!is.na(size) & size > 0){
         detected[i, j] <- rbinom(1, size = size, prob = prob)
       }
    }
  }
  stoch_partial_obs$daily_detected[[k]] <- detected
}

# saveRDS(stoch_partial_obs, paste0("outputs/stoch_I0_100_partial_obs_", tstamp, ".rds"))
```

## Total dataset
```{r}
Rt_dataset <- list()
Rt_dataset$README <- "To see if potential bias of Rt estimation based on the confirmation data set, especially in the most recent time, is reduced or eliminated when the number of cases are large"
Rt_dataset$ode <- out
Rt_dataset$stoch_perfect_obs <- stoch
Rt_dataset$stoch_partial_obs <- stoch_partial_obs
Rt_dataset$Rt <- sapply(0:tend, function(x) get_Rt(x))
Rt_dataset$time <- out$time
Rt_dataset$params <- params
Rt_dataset$y0 <- params$Y0
# saveRDS(Rt_dataset, paste0("outputs/Rt_dataset_I0=100_", tstamp, ".rds"))
```

### Plot: Rt, infection/confirmation/stoch outputs
```{r}
tstamp <- "20211225"
dat <- readRDS(paste0("outputs/Rt_dataset_", tstamp, ".rds"))
# # remove detection data set -- seems wrong
# dat$stoch_partial_obs <- NULL
# saveRDS(dat, paste0("outputs/Rt_dataset_", "20211228", ".rds"))
library(ggplot2)
plt <- ggplot() +
  geom_line(aes(x=dat$time, y = dat$Rt)) +
  labs(x="time", y=expression(R[t]))

ggsave("plots/Rt.png", plt)
library(tidyverse)
oderes <- dat$ode[, c("time", "daily_infected", "daily_confirmed")]
oderes %>% pivot_longer(cols = -time) -> oderes

plt <- ggplot() +
  geom_line(aes(x=dat$time, y = dat$Rt)) +
  labs(x="time", y=expression(R[t]))

ggsave("plots/Rt.png", plt)

stochres <- dat$stoch_perfect_obs$daily_confirmed[,1:20]
stochres$time <-dat$time

stochres %>% pivot_longer(cols = -time) -> stochres

plt <- ggplot() +
  geom_line(data = stochres, aes(time, value, group=name), color="grey50") +
  geom_line(data = oderes, aes(time, value, color=name), size=1.2) + 
  labs(x="time", y="", color="")
   
ggsave("plots/daily_case.png", plt)
```


```{r}
Rt_dataset$stoch_perfect_obs$daily_confirmed$X1[1:20]
Rt_dataset$stoch_partial_obs$daily_detected[[7]]$X1[1:20]

library(ggplot2)
xx <- tend + 1
k <- 2 # sample 

ggplot() +
  geom_line(aes(x=1:xx, y = Rt_dataset$ode$daily_infected)) + 
  geom_line(aes(x=1:xx, y = Rt_dataset$stoch_perfect_obs$daily_infected$X1),
            color = "blue", size = 1, alpha = 0.3)+ 
  geom_line(aes(x=1:xx, y = Rt_dataset$stoch_perfect_obs$daily_infected$X3),
            color = "blue", size = 1, alpha = 0.6)+
  geom_line(aes(x=1:xx, y = Rt_dataset$stoch_partial_obs$daily_detected[[k]]$X1),
            color = "darkred", size = 1, alpha = 0.3)+
  geom_line(aes(x=1:xx, y = Rt_dataset$stoch_partial_obs$daily_detected[[k]]$X3),
            color = "darkred", size = 1, alpha = 0.6)+
  geom_line(aes(x=1:xx, y = Rt_dataset$stoch_partial_obs$daily_detected[[k]]$X1),
            color = "darkred")+
  geom_line(aes(x=1:xx, y = Rt_dataset$stoch_partial_obs$daily_detected[[k]]$X3),
            color = "darkred")+
  labs(x = "day", y = "daily infection")

plot(Rt_dataset$stoch_perfect_obs$daily_confirmed[,1], 
     type = "l", col = rgb(0.6,0.6,0.6, alpha=0.8), ylim=c(0,300))
for (i in 2:nrun){
  lines(Rt_dataset$stoch_perfect_obs$daily_confirmed[,i],
        col = rgb(0.6,0.6,0.6, alpha=0.8))
}
lines(Rt_dataset$ode$daily_confirmed, lwd=2)
# row_mean <- rowMeans(Rt_dataset$stoch_perfect_obs$confirmed, na.rm=TRUE)
# for (i in 1:nrow(daily_case_list$confirmed)){
#   daily_case_list$confirmed[i, is.na(daily_case_list$confirmed[i, ])] <- 0 
# }
row_mean <- rowMeans(Rt_dataset$stoch_perfect_obs$daily_confirmed, na.rm=TRUE)
lines(row_mean, lwd=2, col=2)
```


### Particle filtering - pfilter
```{r}
# library(Kmisc)

tstamp <- "20220105"
I0 <- 10
dat <- readRDS(paste0("outputs/Rt_dataset_I0=", I0, "_", tstamp, ".rds"))

dtype <- "confirmation"

d <- dat$ode[, c("time", "daily_infected")]
if (dtype == "confirmation") {
  d <- dat$ode[, c("time", "daily_confirmed")]
}

dat$params$time_dep_Rt <- FALSE

library(truncnorm)
params <- dat$params

tbegin <- Sys.time()

pf <- pfilter(
  params = params,
  y =  params$Y0,
  data = d,
  data_type = dtype,
  npart = 1e4,
  tend = nrow(d),
  dt = 0.1,
  sd_factor = 4,
  error_pdf = "pois",
  systematic_resampling = FALSE,
  backward_sampling = TRUE,
  stoch = FALSE)

# dat$params[["betavol"]] <- 0.2
# pf <- particle_filter(
#   params = dat$params,
#   y = params$Y0,
#   data = d,
#   data_type = dtype,
#   npart = 1e4,
#   tend = nrow(d),
#   dt = 0.1,
#   error_pdf = "pois",
#   systematic_resampling = FALSE,
#   filter_traj = TRUE,
#   stoch = FALSE)

Sys.time() - tbegin

# Rt <- readRDS("outputs/Rt_default.rds")
epsilon <- params[["epsilon"]]
delta <- params[["delta"]]
gamma <- params[["gamma"]]
bp <- params[["bp"]] # relative infectiousness of pre-symptomatic state
ba <- params[["ba"]] # relative infectiousness of asymptomatic state
fa <- params[["fa"]] # fraction of asymptomatic state

durP <- (1 / delta - 1 / epsilon)
durI <- (1 / gamma)
R0_dur <- ((1 - fa) + fa * ba) * durI + bp * durP

med <- apply(pf$beta_filtered, 1, quantile, probs = c(0.5))
lower <- apply(pf$beta_filtered, 1, quantile, probs = c(0.025))
upper <- apply(pf$beta_filtered, 1, quantile, probs = c(0.975))

# plot(dat$time + 2, dat$daily_Rt, type="l", ylim=c(0, max(upper * R0_dur)))
plot(dat$ode$time + 2, dat$ode$daily_Rt, type="l")
lines(med * R0_dur, lwd=2, col=3)
# lines(upper * R0_dur, lty=2, lwd=2, col=3)
# lines(lower * R0_dur, lty=2, lwd=2, col=3)
```

### Particle filtering - particle_filter
```{r}
# library(Kmisc)

tstamp <- "20220105"
I0 <- 10
dat <- readRDS(paste0("outputs/Rt_dataset_I0=", I0, "_", tstamp, ".rds"))

dtype <- "confirmation"

d <- dat$ode[, c("time", "daily_infected")]
if (dtype == "confirmation") {
  d <- dat$ode[, c("time", "daily_confirmed")]
}

dat$params$time_dep_Rt <- FALSE

library(truncnorm)
params <- dat$params
params[["betavol"]] <- 0.2
tbegin <- Sys.time()


pf <- particle_filter(
  params = params,
  y = params$Y0,
  data = d,
  data_type = dtype,
  npart = 1e4,
  tend = nrow(d),
  dt = 0.1,
  error_pdf = "pois",
  systematic_resampling = FALSE,
  backward_sampling = TRUE,
  stoch = FALSE)

Sys.time() - tbegin

# Rt <- readRDS("outputs/Rt_default.rds")
epsilon <- params[["epsilon"]]
delta <- params[["delta"]]
gamma <- params[["gamma"]]
bp <- params[["bp"]] # relative infectiousness of pre-symptomatic state
ba <- params[["ba"]] # relative infectiousness of asymptomatic state
fa <- params[["fa"]] # fraction of asymptomatic state

durP <- (1 / delta - 1 / epsilon)
durI <- (1 / gamma)
R0_dur <- ((1 - fa) + fa * ba) * durI + bp * durP

med <- apply(pf$beta_filtered, 1, quantile, probs = c(0.5))
lower <- apply(pf$beta_filtered, 1, quantile, probs = c(0.025))
upper <- apply(pf$beta_filtered, 1, quantile, probs = c(0.975))

# plot(dat$time + 2, dat$daily_Rt, type="l", ylim=c(0, max(upper * R0_dur)))
plot(dat$ode$time + 2, dat$ode$daily_Rt, type="l")
lines(med * R0_dur, lwd=2, col=3)
# lines(upper * R0_dur, lty=2, lwd=2, col=3)
# lines(lower * R0_dur, lty=2, lwd=2, col=3)
```

### Backward sampling: extract_trace - ODE resutls
```{r}
# devtools::load_all()
# devtools::install()
# library(Kmisc)
library(ggplot2)
theme_set(theme_bw())

tstamp <- "20220105"
dat <- readRDS(paste0("outputs/Rt_dataset_I0=10_", tstamp, ".rds"))
params <- dat$params
params$time_dep_Rt <- FALSE

nrep <- 1e2
npart <- 1e4
dt <- 0.1
library(parallel)
library(doParallel)
library(foreach)

ncores <- detectCores() 
data_type <- "confirmation"
# ODE with perfect observation
d <- data.frame(date = dat$time,
               daily_infected = round(dat$ode$daily_infected))
# 
if (data_type == "confirmation") {
  d <- data.frame(date = dat$time,
                  daily_confirmed = round(dat$ode$daily_confirmed))
}

set.seed(42)
cl <- makeCluster(getOption("cl.cores", 2))
doParallel::registerDoParallel(cl)
  
pf <- foreach(i = 1:nrep, .packages = c("Kmisc"), .inorder = F) %dopar% {
  extract_trace(params = dat$params,
                y = dat$params$Y0,
                data = d,
                data_type = data_type,
                npart = npart,
                tend = nrow(d),
                dt = 0.1,
                sd_factor = 5,
                error_pdf = "pois",
                negbin_size = 15,
                backward_sampling = TRUE,
                stoch = FALSE)
}
parallel::stopCluster(cl)
  
parset_chr <- 
    paste0(data_type, "_I0=", params$I0, "_npart=", npart,
           "_nrep=", nrep, "_dt=", dt, "_", tstamp)
  
# saveRDS(pf, paste0("outputs/pf_", parset_chr, ".rds"))
  
pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
Rt_est <- as.data.frame(sapply(pf, function(x) x[, "Rt"]))
Rt_quantile <- as.data.frame(t(apply(Rt_est, 1, function(x) quantile(x, pr))))
  
df <- cbind(Rt_quantile, d)
col_fill <- "#1F618D"
col_dat <- "grey70"
df$Rt <- dat$ode$daily_Rt

plt <- ggplot(df, aes(x = date)) +
    geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = col_fill, alpha = 0.5) +
    geom_ribbon(aes(ymin = `25%`, ymax = `75%`), fill = col_fill, alpha = 0.8) +
    geom_line(aes(y = `50%`), color = col_fill, size = 1.2) +
    geom_line(aes(x = date + 1, y = Rt), size = 1, linetype = "dashed") +
    geom_hline(yintercept = 1, color = "darkred", size = 1, linetype = "dotted")+
    labs(title = parset_chr, y = expression(italic(R)[italic(t)]), x = "")

plt  
# ggsave(paste0("plots/", parset_chr, ".png"), plt)

```

### Backward sampling: extract_trace2 - backward_sample parameter in the function
works file
```{r}
# devtools::load_all()
# devtools::install()
# library(Kmisc)

library(ggplot2)
theme_set(theme_bw())

tstamp <- "20220105"
I0 <- 100
dat <- readRDS(paste0("outputs/Rt_dataset_I0=", I0, "_", tstamp, ".rds"))
params <- dat$params
params$time_dep_Rt <- FALSE

nrep <- 1e3
npart <- 1e4
dt <- 0.2
library(parallel)
library(doParallel)
library(foreach)

ncores <- detectCores() 
data_type <- "confirmation"
# ODE with perfect observation
# d <- data.frame(date = dat$time,
#                daily_infected = round(dat$ode$daily_infected))
# # 
# if (data_type == "confirmation") {
#   d <- data.frame(date = dat$time,
#                   daily_confirmed = round(dat$ode$daily_confirmed))
# }
# 
for (k in 13:100) {
  d <- data.frame(date = dat$time,
                 daily_confirmed = round(dat$stoch_perfect_obs$daily_confirmed[,k]))
  
  tailsum <- sum(tail(d$daily_confirmed))
  message(paste0("k = ", k, ", tailsum = ", tailsum))
  
  if (tailsum > 0) {
    tbegin <- Sys.time()
    
    set.seed(42)
    cl <- makeCluster(getOption("cl.cores", 2))
    doParallel::registerDoParallel(cl)
    
    params[["betavol"]] <- 0.3
    
    pf <- foreach(i = 1:nrep, .packages = c("Kmisc"), .inorder = F) %dopar% {
      extract_trace2(params = params,
                    y = params$Y0,
                    data = d,
                    data_type = data_type,
                    npart = npart,
                    tend = nrow(d),
                    dt = dt,
                    error_pdf = "pois",
                    negbin_size = 15,
                    backward_sampling = TRUE,
                    stoch = FALSE)
    }
    parallel::stopCluster(cl)
      
    telapsed <- Sys.time() - tbegin
    message(paste0("time elapsed = ", telapsed))
    
    parset_chr <- 
        paste0(data_type, "_I0=", params$I0, "_betavol=", params[["betavol"]], "_npart=", npart, "_nrep=", nrep, "_dt=", dt, "_", "k=", k, "_", tstamp)
      
    saveRDS(pf, paste0("outputs/pf_", parset_chr, ".rds"))
      
    pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
    Rt_est <- as.data.frame(sapply(pf, function(x) x[, "Rt"]))
    Rt_quantile <- as.data.frame(t(apply(Rt_est, 1, function(x) quantile(x, pr))))
      
    df <- cbind(Rt_quantile, d)
    col_fill <- "#1F618D"
    col_dat <- "grey70"
    df$Rt <- dat$ode$daily_Rt
    
    plt <- ggplot(df, aes(x = date)) +
        geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = col_fill, alpha = 0.5) +
        geom_ribbon(aes(ymin = `25%`, ymax = `75%`), fill = col_fill, alpha = 0.8) +
        geom_line(aes(y = `50%`), color = col_fill, size = 1) +
        geom_line(aes(x = date + 1, y = Rt), size = 1, linetype = "dashed") +
        geom_hline(yintercept = 1, color = "darkred", size = 1, linetype = "dotted")+
        labs(title = parset_chr, y = expression(italic(R)[italic(t)]), x = "")
        # scale_y_continuous(limits=c(0,3))
    
    # plt  
    ggsave(paste0("plots/", parset_chr, ".png"), plt)
  }
}
```

### Backward sampling using rolling average
```{r}
# devtools::load_all()
# devtools::install()
# library(Kmisc)

library(ggplot2)
theme_set(theme_bw())

tstamp <- "20220105"
I0 <- 100
dat <- readRDS(paste0("outputs/Rt_dataset_I0=", I0, "_", tstamp, ".rds"))
params <- dat$params
params$time_dep_Rt <- FALSE

nrep <- 1e3
npart <- 1e4
dt <- 0.2
library(parallel)
library(doParallel)
library(foreach)

ncores <- detectCores() 
data_type <- "confirmation"
# ODE with perfect observation
# d <- data.frame(date = dat$time,
#                daily_infected = round(dat$ode$daily_infected))
# # 
# if (data_type == "confirmation") {
#   d <- data.frame(date = dat$time,
#                   daily_confirmed = round(dat$ode$daily_confirmed))
# }
# 
roll_window <- 7
for (k in 1:20) {
  dd <- dat$stoch_perfect_obs$daily_confirmed[,k]
  roll_case <- 
    data.table::frollmean(dd, n=roll_window, align="right")
           
  roll_case[1:6] <- dd[1:6] # include original data                          
  
  d <- data.frame(date = dat$time,
                 daily_confirmed = round(roll_case))
  
  tailsum <- sum(tail(d$daily_confirmed))
  message(paste0("k = ", k, ", tailsum = ", tailsum))
  
  if (tailsum > 0) {
    tbegin <- Sys.time()
    
    set.seed(42)
    cl <- makeCluster(getOption("cl.cores", 2))
    doParallel::registerDoParallel(cl)
    
    params[["betavol"]] <- 0.3
    
    pf <- foreach(i = 1:nrep, .packages = c("Kmisc"), .inorder = F) %dopar% {
      extract_trace2(params = params,
                    y = params$Y0,
                    data = d,
                    data_type = data_type,
                    npart = npart,
                    tend = nrow(d),
                    dt = dt,
                    error_pdf = "pois",
                    negbin_size = 15,
                    backward_sampling = TRUE,
                    stoch = FALSE)
    }
    parallel::stopCluster(cl)
      
    telapsed <- Sys.time() - tbegin
    message(paste0("time elapsed = ", telapsed))
    
    parset_chr <- 
        paste0(data_type, "_I0=", params$I0, "_betavol=", params[["betavol"]], "_npart=", npart, "_nrep=", nrep, "_dt=", dt, "_", "k=", k, "_rollwd=", roll_window, "_", tstamp)
      
    saveRDS(pf, paste0("outputs/pf_", parset_chr, ".rds"))
      
    pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
    Rt_est <- as.data.frame(sapply(pf, function(x) x[, "Rt"]))
    Rt_quantile <- as.data.frame(t(apply(Rt_est, 1, function(x) quantile(x, pr))))
      
    df <- cbind(Rt_quantile, d)
    col_fill <- "#1F618D"
    col_dat <- "grey70"
    df$Rt <- dat$ode$daily_Rt
    
    plt <- ggplot(df, aes(x = date)) +
        geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = col_fill, alpha = 0.5) +
        geom_ribbon(aes(ymin = `25%`, ymax = `75%`), fill = col_fill, alpha = 0.8) +
        geom_line(aes(y = `50%`), color = col_fill, size = 1) +
        geom_line(aes(x = date + 1, y = Rt), size = 1, linetype = "dashed") +
        geom_hline(yintercept = 1, color = "darkred", size = 1, linetype = "dotted")+
        labs(title = parset_chr, y = expression(italic(R)[italic(t)]), x = "")
        # scale_y_continuous(limits=c(0,3))
    
    # plt  
    ggsave(paste0("plots/", parset_chr, ".png"), plt)
  }
}  
```



### Fit - stochastic model resutls
```{r}
tstamp <- "20211225"
dat <- readRDS(paste0("outputs/Rt_dataset_", tstamp, ".rds"))
library(Kmisc)
library(pfilter)
# devtools::install()
nrep <- 1e4
npart <- 1e4
dt <- 0.2

library(ggplot2)
theme_set(theme_bw())
library(parallel)
library(doParallel)
library(foreach)

ncores <- detectCores() 

theta["betavol"] <- 0.25

for (id in 1:10) {
  d <- data.frame(date = dat$time,
                daily_confirmed = dat$stoch_perfect_obs$daily_confirmed[,id])
  
  parset_chr <- 
    paste0("stoch_ID=", id, "_npart=", npart, "_nrep=", nrep, "_dt=", dt)
    
  set.seed(42)

  cl <- makeCluster(getOption("cl.cores", 2))
  doParallel::registerDoParallel(cl)
  
  pf <- foreach(i = 1:nrep, .packages = c("pfilter"), .inorder = F) 
  %dopar% {
    extract_trace(params = theta,
                  y = y0,
                  data = d,
                  data_type = "confirmation",
                  npart = npart,
                  tend = nrow(d),
                  dt = dt,
                  error_pdf = "pois",
                  negbin_size = 15,
                  stoch = FALSE)
  }
  parallel::stopCluster(cl)
  
  saveRDS(pf, paste0("outputs/pf_", tstamp, "_", parset_chr, ".rds"))
  
  pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
  Rt_est <- as.data.frame(sapply(pf, function(x) x[, "Rt"]))
  Rt_quantile <- as.data.frame(t(apply(Rt_est, 1, function(x) quantile(x, pr))))
  
  df <- cbind(Rt_quantile, d)
  col_fill <- "#1F618D"
  col_dat <- "grey70"
  df$Rt <- dat$Rt
  
  plt <- ggplot(df, aes(x = date)) +
    geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = col_fill, alpha = 0.5) +
    geom_ribbon(aes(ymin = `25%`, ymax = `75%`), fill = col_fill, alpha = 0.8) +
    geom_line(aes(y = `50%`), color = col_fill, size = 1.2) +
    geom_line(aes(y = Rt), size = 1, linetype = "dashed") +
    geom_hline(yintercept = 1, color = "darkred", size = 1, linetype = "dotted")+
    labs(title = parset_chr, y = expression(italic(R)[italic(t)]), x = "")
  
  ggsave(paste0("plots/", parset_chr, ".png"), plt)
}
```


### Fit - stochastic simulation results
```{r}
## forward filtering with backward sampling
# d <- out[1:175, c("time", "daily_confirmed")]
# names(d) <- c("date", "daily_confirmed")

tstamp <- "20211225"
dat <- readRDS(paste0("outputs/Rt_dataset_", tstamp, ".rds"))
library(Kmisc)
library(pfilter)
# devtools::install()
theta["betavol"] <- 0.25
library(ggplot2)
theme_set(theme_bw())
library(parallel)
library(doParallel)
library(foreach)
set.seed(42)

ncores <- detectCores()

for(id in 1:10){
  d <- data.frame(date = dat$time, 
                daily_confirmed = dat$stoch_perfect_obs$daily_confirmed[, id])


cl <- makeCluster(getOption("cl.cores", ncores - 2))
doParallel::registerDoParallel(cl)

pf <- foreach(i = 1:1e3, .packages = c("pfilter"), .inorder = F) %dopar% {
  extract_trace(params = theta,
                y = y0,
                data = d,
                data_type = "confirmation",
                npart = 1e4,
                tend = nrow(d),
                dt = 0.2,
                error_pdf = "pois",
                negbin_size = 15,
                stoch = FALSE)
}
parallel::stopCluster(cl)

pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
Rt_est <- as.data.frame(sapply(pf, function(x) x[, "Rt"]))
Rt_quantile <- as.data.frame(t(apply(Rt_est, 1, function(x) quantile(x, pr))))

df <- cbind(Rt_quantile, d)
col_fill <- "#1F618D"
col_dat <- "grey70"
df$Rt <- dat$Rt
plt1 <- ggplot(df, aes(x = date)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = col_fill, alpha = 0.5) +
  geom_ribbon(aes(ymin = `25%`, ymax = `75%`), fill = col_fill, alpha = 0.8) +
  geom_line(aes(y = `50%`), color = col_fill, size = 1.2) +
  geom_line(aes(y = Rt), size = 1, linetype = "dashed") +
  geom_hline(yintercept = 1, color = "darkred", size = 1, linetype = "dotted")+
  labs(title = paste0("ID = ", id, ", dt = 0.1"), y = expression(italic(R)[italic(t)]), x = "")
  # 
  # labs(title = paste0("ID = ", id, ", Prob = 0.8"), y = expression(italic(R)[italic(t)]), x = "")
  
  # labs(title = expression(italic(R)[italic(t)]~estimated~using~particle~filtering),
  #        y = expression(italic(R)[italic(t)]), x = "") 

# ggsave(paste0("plots/pf_stoch_confirmed_Rt_", id, ".png"), plt)

daily_conf <- as.data.frame(sapply(pf, function(x) x[, "CR"]))
daily_conf_quantile <- 
  as.data.frame(t(apply(daily_conf, 1, function(x) quantile(x, pr))))
df <- cbind(daily_conf_quantile, d)
df$ode_traj <- dat$ode$daily_confirmed
# df$ode_traj <- dat$ode$daily_infected
plt2 <- ggplot(df, aes(x = date)) +
  geom_col(aes(y = daily_confirmed), color = col_dat, width = 0.8) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = col_fill, alpha = 0.4) +
  geom_ribbon(aes(ymin = `25%`, ymax = `75%`), fill = col_fill, alpha = 0.6) +
  geom_line(aes(y = `50%`), color = col_fill, size = 1.2) +
  geom_line(aes(y = ode_traj), size = 1, linetype = "dashed")+
  labs(x = "", y = "Daily confirmed case") 
  
# ggsave(paste0("plots/pf_stoch_confirmed_case_", id, ".png"), plt)

gA <- ggplotGrob(plt1)
gB <- ggplotGrob(plt2)
png(paste0("plots/pf_stoch_confirm_id_", id, "_", tstamp, ".png"),
    width = 480*1.4, height = 480*1.4, res = 1.4*72)
grid::grid.newpage()
grid::grid.draw(rbind(gA, gB))
dev.off()

# plts <- gridExtra::grid.arrange(plt1, plt2, nrow = 2)
# ggsave(paste0("plots/proc_stoch_pf_stoch__", id, ".png"), plts)
}
```


## Simulate using the estimated $R_t$ 
```{r}
library(Kmisc)
# extract paramter (str) of interest from particle filter results (pf)
# pf is a list of data.frame where row is time and column is sample
# days is a number of rows to extract. i.e., days = 7 leads the model to 
# calculate the mean Rt from the last 7 days for number of samples
npart <- 1e4
nsample <- 200
ids <- sample.int(npart, nsample)

extract_sample_pf <- function(pf, str = "Rt", days = 1, ids = NULL){
  df <- as.data.frame(sapply(pf, function(x) x[, str]))
  nr <- nrow(df)
  df_sub <- df[(nr - days + 1):nr, ids]
  if(days > 1) {
   vec <- colMeans(df_sub)
  }
  else{
    vec <- unlist(df_sub)
  } 
  return (vec)
}

pf <- readRDS("outputs/pf_20220102_confirm_20220102_betavol=0.3_npart=10000_nrep=10000_dt=0.1.rds")

str <- names(pf[[1]])
samp <- lapply(str[1:9], function(x) extract_sample_pf(pf, x, 1, ids))
samp_Rt <- lapply(str[10], function(x) extract_sample_pf(pf, x, 1, ids))

tend <- 30
tstamp <- "20220102" #"20211225"
dat <- readRDS(paste0("outputs/Rt_dataset_", tstamp, ".rds"))
params <- as.list(dat$params)
params$tend <- tend

sub_pf <- pf[ids]
fit_confirm <- as.data.frame(sapply(sub_pf, function(x) x[, "CR"]))

# add 1 to put the last data points 
proj_confirm <- data.frame(matrix(NA, nrow = tend + 1, ncol = nsample))
proj_confirm[1, ] <- tail(fit_confirm, 1)
library(deSolve)
library(dplyr)
for (i in 1:nsample) {
  y <- sapply(samp, function(x) x[i])
  names(y) <- names(THETA$Y0)
  params$R0 <- samp_Rt[i]
  params$time_dep_Rt <- FALSE
  ode(y = yy, times = seq(0, tend, by = 1), func = ode_sepair, 
             parms = params, Rt_varying = FALSE) %>%
    as.data.frame() -> res
  proj_confirm[2:nrow(proj_confirm), i] <- diff(res$CR)
}

pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
fit_confirm_qt <- as.data.frame(t(apply(fit_confirm, 1, function(x) quantile(x, pr))))
proj_confirm_qt <- as.data.frame(t(apply(proj_confirm, 1, function(x) quantile(x, pr))))

sub1 <- data.frame(matrix(NA,
                   nrow = nrow(proj_confirm_qt) - 1,
                   ncol = ncol(proj_confirm_qt)))

names(sub1) <- names(fit_confirm_qt) 
  
df1 <- rbind(fit_confirm_qt, sub1)

sub2 <- data.frame(matrix(NA,
                   nrow = nrow(fit_confirm_qt) - 1, # -1 to match the end of the fit
                   ncol = ncol(fit_confirm_qt)))
names(sub2) <- names(fit_confirm_qt) 

df2 <- rbind(sub2, proj_confirm_qt)
df1$time <- 0:(nrow(df1)-1)
df2$time <- 0:(nrow(df2)-1)

library(ggplot2)
ggplot(df1, aes(x = time)) +
  geom_col(data = dat$ode, aes(x = time, y = daily_confirmed), fill = "grey50", alpha = 0.3) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), 
              fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = `25%`, ymax = `75%`), 
              fill = "steelblue", alpha = 0.7) +
  geom_line(aes(y = `50%`), color = "steelblue", size = 1) + 
  geom_ribbon(data = df2, aes(x = time, ymin = `2.5%`, ymax = `97.5%`), 
              fill = "indianred", alpha = 0.3) +
  geom_ribbon(data = df2, aes(x = time, ymin = `25%`, ymax = `75%`), 
              fill = "indianred", alpha = 0.7) +
  geom_line(data = df2, aes(x = time, y = `50%`), color = "indianred", size = 1)
```

### Run tauleap model
```{r}
nsample <- 10000
tend <- 30
day_confirm_tau <- data.frame(matrix(NA, nrow = nsample, ncol = tend))

for (i in 1:nsample) {
val <- sapply(samp, function(x) x[1,i])
  yy <- val[1:8]
  names(yy) <- names(y0)
  params$R0 <- val[9]
  res_tauleap <- solve_tauleap(sepair_tauleap, 
                             y = round(yy), 
                             times = 0:tend, 
                             params = params,
                             delta = 0.2) 
  day_confirm_tau[i,] <- diff(res_tauleap$R)
}

plot(1:tend, day_confirm_tau[1,], type="l", col="#F5927D",
     ylim=c(min(day_confirm_tau),max(day_confirm_tau)))
for (i in 2:nsample) {
  lines(1:tend, day_confirm_tau[i,], col="#F5927D")
}

lines(1:tend, colMeans(day_confirm_tau), lwd=2, col=1)

# res_tauleap <- solve_tauleap(sepair_tauleap, 
#                              y = round(params$y0), 
#                              times = 0:params$tend, 
#                              params = params,
#                              delta = 0.2) 
#   
# lines(1:tend, diff(res_tauleap$R), col="#F5927D")

df <- data.frame(t=1:100, A=c(rnorm(80),rep(NA,20)), B=c(rep(NA,80),rpois(20,lambda=2)))
library(ggplot2)
ggplot(df) +
  geom_line(aes(t,A), color = "steelblue") + 
  geom_line(aes(t,B), color = "indianred") 
```



### 
```{r}
pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
pf <- readRDS("outputs/pf_20211225_betavol=0.25_npart=10000_nrep=10000_dt=0.1.rds")
df <- as.data.frame(sapply(pf, function(x) x[, "Rt"]))
Rt_quantile <- as.data.frame(t(apply(df, 1, function(x) quantile(x, pr))))

# extract state variables
str_list <- list("S", "E", "P", "A", "I", "R", "Rt")
sapply(str_list, function(x) sample_values_pf(pf, str = x, n = 1, nsample = 200))

df <- cbind(Rt_quantile, dat[, c("date", "daily_confirmed")])
col_fill <- "#1F618D"
# col_dat <- "#F1948A"
col_dat <- "grey80"
# col_rt <- "#148F77"
col_rt <- "#1F618D"
plt <- ggplot(df, aes(x = date)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = col_rt, alpha = 0.3) +
  geom_ribbon(aes(ymin = `25%`, ymax = `75%`), fill = col_rt, alpha = 0.7) +
  geom_line(aes(y = `50%`), color = col_rt, size = 1) + 
  geom_hline(yintercept = 1, color = "darkred", size = 1, linetype = "dotted")+
  labs(title = expression(R[t]~estimated~using~particle~filtering),
       y = expression(R[t]), x = "") +
  scale_x_date(date_breaks = "1 month")

ggsave(paste0("plots/Rt", tstamp, ".png"), plt, 
       width = 8, height = 5,
       dpi = 600, scale = 2, units = "cm", type = "cairo")
```

## EpiNow2
```{r}
Rt <- readRDS("outputs/Rt_dataset_20210809T215134.rds")
library(EpiNow2)
# set number of cores to use
options(mc.cores = ifelse(interactive(), 4, 1))
# get example case counts
reported_cases <- example_confirmed[1:60]

# set up example generation time
generation_time <- get_generation_time(disease = "SARS-CoV-2", source = "ganyani")
# set delays between infection and case report 
incubation_period <- get_incubation_period(disease = "SARS-CoV-2", source = "lauer")
reporting_delay <- list(mean = convert_to_logmean(3, 1), mean_sd = 0.1,
                        sd = convert_to_logsd(3, 1), sd_sd = 0.1, max = 10)
      
# default setting 
# here we assume that the observed data is truncated by the same delay as 
def <- estimate_infections(reported_cases, generation_time = generation_time,
                           delays = delay_opts(incubation_period, reporting_delay),
                           rt = rt_opts(prior = list(mean = 2, sd = 0.1)),
                           stan = stan_opts(control = list(adapt_delta = 0.95)))
# real time estimates
summary(def)
# summary plot
plot(def)
```

```{r}
Rt <- readRDS("outputs/Rt_dataset_20210809T215134.rds")
reported_cases <- data.frame(date = seq(as.Date("2020-02-22"), length.out = length(Rt$stoch_perfect_obs$confirmed[[1]]), by = "day"), 
                             confirm =  Rt$stoch_perfect_obs$confirmed[[1]])

# set up example generation time
generation_time$mean <- 6.5
# set delays between infection and case report 
incubation_period$mean <- 5
reporting_delay$mean <- 2.5

# default setting 
# here we assume that the observed data is truncated by the same delay as 
def <- estimate_infections(reported_cases, generation_time = generation_time,
                           delays = delay_opts(incubation_period, reporting_delay),
                           rt = rt_opts(prior = list(mean = 2, sd = 0.1)),
                           stan = stan_opts(control = list(adapt_delta = 0.95)))
# real time estimates
summary(def)
# summary plot
plot(def)
```

