---
title: "pomp package"
output: html_notebook
editor_options: 
  chunk_output_type: console
---



## fake data generation
```{r}
library(Kmisc)
library(deSolve)
library(tidyverse)
devtools::load_all()
params <- c(beta = 0.3, gamma = 0.2)
y0 <- c(S = 9950, I = 50, R = 0, CI = 0)
tend <- 100
times <- seq(0, tend, by = 1) # daily output for 150 days

ode(y = y0, times = times, func = ode_sir, parms = params) %>% 
  as.data.frame() -> out 

out$daily_symptom_onset <- c(0, diff(out$CI))
set.seed(23)
# out$obs <- rpois(tend+1, out$daily_symptom_onset)
out$obs <- round(out$daily_symptom_onset)
```

## ODE model
### ODE model compile and run
```{r}
library(pomp)
sir_ode <- "
  double N = S + I + R;
  double infection = S*beta*I/N;
  double recovery = I*gamma;
  DS = -infection;
  DI = infection - recovery;
  DR = recovery;
  DY = infection; 
"
rmeas <-"
  Z = rpois(Y);
"
dmeas <-"
  lik = dpois(Z, Y, give_log);
"
rinit <-"
  S=9950;
  I=50;
  R=0;
  Y=0;
"

pomp(data = data.frame(time=seq(0,100,by=1),Z=out[, "obs"]),
     times = "time",
     t0 = 0,
     rmeasure = Csnippet(rmeas),
     dmeasure = Csnippet(dmeas),
     skeleton = vectorfield(Csnippet(sir_ode)),
     statenames = c("S","I","R","Y"),
     paramnames = c("beta","gamma"),
     params = c(beta=0.2,gamma=0.1),
     accumvars = "Y",
     rinit = Csnippet(rinit)) -> sirodeC 

sirodeC %>%
  trajectory(params = c(beta=0.3, gamma=0.2),
             format="data.frame") -> res
plot(res)

p <- parmat(c(beta=0.3,gamma=0.2),10)
p["beta",] <- seq(0.1,0.4,length=10)
sirodeC %>%
  trajectory(params=p,format="data.frame") %>%
  ggplot(mapping=aes(x=time,y=I,color=.id,group=.id))+
  guides(color="none")+
  geom_line()+
  theme_bw()
```

### ODE Parameter estimation -- trajectory matching 
```{r}
sirodeC %>%
  traj_objfun(est = c("beta","gamma"),
              params = c(beta=0.5,gamma=0.3),
              dmeasure = Csnippet(dmeas),
              statenames = c("S","I","R","Y"),
              paramnames = c("beta","gamma")) -> ofun


ofun(c(0.3,0.2))
library(subplex)
subplex(c(beta=0.2,gamma=0.15),fn=ofun) -> fit
fit
ofun(fit)
coef(fit)
logLik(fit)

ofun %>%
  trajectory(format="data.frame") %>%
  ggplot(mapping=aes(x=time,y=Y,color=.id,group=.id))+
  geom_point(data=out, aes(x=time,y=daily_symptom_onset), color="darkred", inherit.aes=F)+
  guides(color="none")+
  geom_line()+
  theme_bw()

sirodeC %>%
  traj_objfun(
    est=c("beta","gamma"),
    params=c(beta=0.5,gamma=0.1),
    partrans=parameter_trans(log=c("beta","gamma")),
    dmeasure=Csnippet(dmeas), 
    statenames=c("S","I","R","Y"), 
    paramnames=c("beta","gamma")) -> ofun2

subplex(log(c(beta=0.1,gamma=0.2)),fn=ofun2) -> fit
fit
ofun2(fit)
ofun2(fit$par)
coef(fit$par)
fit$value
ofun2(fit$par)
```

## Stochastic model
```{r}
library(pomp)
sir_step <- "
  double N = S + I + R;
  double lambda = beta * I / N;
  double infection = rbinom(S, 1-exp(-lambda*dt));
  double recovery = rbinom(I, 1-exp(-gamma*dt));
  S += -infection;
  I += infection - recovery;
  R += recovery;
  Y += infection; 
"

rmeas <-"
  Z = rpois(Y);
"

dmeas <-"
  lik = dpois(Z, Y, give_log);
"

rinit <-"
  S=9950;
  I=50;
  R=0;
  Y=0;
"

pomp(data = data.frame(time=seq(0,100,by=1),Z=out[, "obs"]),
     times = "time",
     t0 = 0,
     rmeasure = Csnippet(rmeas),
     dmeasure = Csnippet(dmeas),
     rprocess = euler(Csnippet(sir_step), delta.t = 0.2),
     statenames = c("S","I","R","Y"),
     paramnames = c("beta","gamma"),
     accumvars = "Y",
     rinit = Csnippet(rinit)) -> sir

sir %>%
  simulate(params=c(beta=0.3,gamma=0.2)) %>%
  plot()

# # multiple simulation
# simulate(sir, params=c(beta=0.3,gamma=0.2), nsim = 100) -> sims
# ggplot(data=gather(as.data.frame(sims), variable, value, Y, Z),
#        aes(x = time, y = value, color = variable,
#            group = interaction(.id, variable)))+
#   geom_line()+
#   facet_grid(variable~.,scales="free_y")+
#   labs(y="",color="")
```

## Parameter estimation using iterated filtering (IF) - local search
```{r}
# Local search
set.seed(234)
sir %>%
  mif2(
    params=c(beta=0.5, gamma=0.5),
    Np=1000,
    Nmif=50,
    partrans=parameter_trans(log=c("beta","gamma")),
    paramnames=c("beta","gamma"),
    rw.sd=rw.sd(beta=0.01,gamma=0.01),
    cooling.fraction.50=0.5) -> mf1

coef(mf1)
plot(mf1)
```

# Global search - IF from multiple initial points
```{r}
# Global search
sobol_design(
  lower = c(beta = 0.01, gamma = 0.01),
  upper = c(beta = 0.99, gamma = 0.99),
  nseq = 10
) -> guesses
# guesses

library(parallel)
library(doParallel)
library(foreach)

ncores <- detectCores()
cl <- makeCluster(getOption("cl.cores", ncores - 2))
doParallel::registerDoParallel(cl)

foreach (guess = iter(guesses, "row"),
         .combine = c,
         .export = c("guesses"),
         .packages = c("pomp"),
         # .errorhandling = "remove",
         .inorder = FALSE) %dopar% {mf1 %>% mif2(params = guess)} -> mifs

parallel::stopCluster(cl)

mifs %>%
  traces() %>%
  melt() %>%
  ggplot(aes(x=iteration,y=value,group=L1,color=L1))+
  geom_line()+
  facet_wrap(~variable,scales="free_y")+
  guides(color="none")

mifs %>% 
  as("data.frame") %>% 
  gather(variable,value,-time,-.id) %>%
  ggplot(aes(x=time,y=value,group=.id,color=.id))+
  geom_line()+
  facet_wrap(~variable,scales="free_y",ncol=1)+
  guides(color="none")
```


Why can't we recover the values of beta and gamma from which we started??
When population size is small, round-off errors matter. When the initial
conditions are S=995 and I=5, Y 


# Calculate likelihood    
```{r}
sir %>%
  pfilter(Np=1000,params=c(beta=0.3,gamma=0.2)) -> pfsir

replicate(10, pfsir %>% pfilter() %>% logLik()) -> lls
lls


foreach (mf=mifs,
  .combine=rbind, .packages=c("pomp"), 
  .errorhandling="remove", .inorder=FALSE) %dopar% {
    
    replicate(5, 
      mf %>% pfilter() %>% logLik()
    ) %>%
      logmeanexp(se=TRUE) -> ll
    
    data.frame(as.list(coef(mf)),loglik=ll[1],loglik.se=ll[2])
    
  } -> estimates


```



```{r}
logmeanexp(lls,se=TRUE) -> ll_sir1
ll_sir1

replicate(5, mf1 %>% pfilter() %>% logLik()) %>% logmeanexp(se=TRUE)

sir %>% simulate(nsim=1e2,params=c(beta=coef(mf1)[[1]],gamma=coef(mf1)[[2]])) -> sirsim

## compare fits with the data
plot(sirsim[[1]]@states["Y",], type="l", ylim=c(0,30))
for(i in 2:100)  lines(sirsim[[i]]@states["Y",])
points(out$obs, col=2, lwd=2)
points(sirsim[[1]]@data[1,], col=3)
points(sirsim[[2]]@data[1,], col=3)
for(i in 2:100) points(sirsim[[i]]@data[1,], col=2, lwd=2)

lines(sirsim[[3]]@data[1,], col=2, lwd=2)

```


## $\beta$ as time-dependent variable
```{r}
library(pomp)
sir_step <- "
  logbeta += rnorm(0, vol);
  double N = S + I + R;
  double lambda = exp(logbeta) * I / N;
  double infection = rbinom(S, 1-exp(-lambda*dt));
  double recovery = rbinom(I, 1-exp(-gamma*dt));
  
  S += -infection;
  I += infection - recovery;
  R += recovery;
  Y += infection; 
"

rmeas <-"
  Z = rpois(Y);
"

dmeas <-"
  lik = dpois(Z, Y, give_log);
"

rinit <-"
  S = 9950;
  I = 50;
  R = 0;
  Y = 0;
  logbeta = log(0.3);
"

d <- data.frame(time = seq(0, 100, by=1), Z=out[, "obs"])
pomp(data = d,
     times = "time",
     t0 = 0,
     rmeasure = Csnippet(rmeas),
     dmeasure = Csnippet(dmeas),
     rprocess = euler(Csnippet(sir_step), delta.t = 0.2),
     statenames = c("S", "I", "R", "Y", "logbeta"),
     paramnames = c("vol", "gamma"),
     accumvars = "Y",
     rinit = Csnippet(rinit)) -> sir

sir %>%
  simulate(params=c(gamma=0.2, vol=0.2)) %>%
  plot()

# Local search
set.seed(234)
sir %>%
  mif2(
    params=c(vol=0.5, gamma=0.5),
    Np=1000,
    Nmif=50,
    partrans=parameter_trans(log=c("vol","gamma")),
    paramnames=c("vol","gamma"),
    rw.sd=rw.sd(vol=0.01,gamma=0.01),
    cooling.fraction.50=0.5) -> mf1

coef(mf1)
plot(mf1)


sirsim <- sir %>% 
  simulate(nsim = 1e2, 
           params = c(vol=coef(mf1)[[1]],gamma=coef(mf1)[[2]])) 

## compare fits with the data
plot(sirsim[[1]]@states["Y",], type="l", ylim=c(0,200))
for(i in 2:100)  lines(sirsim[[i]]@states["Y",])
points(out$obs, col=2, lwd=2)
points(sirsim[[1]]@data[1,], col=3)



## beta
plot(exp(sirsim[[1]]@states["logbeta",]), type="l", ylim=c(0,1))
for(i in 2:100)  lines(exp(sirsim[[i]]@states["logbeta",]))

Rt <- data.frame(matrix(NA, nrow=101, ncol=100))
for(i in 1:100) {
  N <- sirsim[[i]]@states["S",]+                                      sirsim[[i]]@states["I",]+sirsim[[i]]@states["R",]
  
  Rt[,i] <- 
    exp(sirsim[[i]]@states["logbeta",])/0.2*sirsim[[i]]@states["S",]/N
}
Rt$rowmean <- rowMeans(Rt)

Rt$t <- 0:100
plot(Rt$t, Rt$rowmean, type="l")
```

## SEPIAR $\beta$ as time-dependent variable
```{r}
library(pomp)
sepiar_step <- Csnippet("
  logbeta += rnorm(0, vol);
  double N = S + E + P + I + A + R;
  double lambda = exp(logbeta) * (I + ba*A + bp*P) / N;
  double infection = rbinom(S, 1-exp(-lambda*dt));
  double E_to_P = rbinom(E, 1-exp(-epsilon*dt));
  double rate = 1/(1/delta - 1/epsilon);
  double rate_from_P[2], trans_from_P[2];
  rate_from_P[0] = rate*(1-fa);
  rate_from_P[1] = rate*fa;

  // Rprintf(\"rate_from_P[0] = %lg \\n\", rate_from_P[0]);
  // Rprintf(\"rate_from_P[1] = %lg \\n\", rate_from_P[1]);
  
  reulermultinom(2, P, &rate_from_P[0], dt, &trans_from_P[0]);
  double recovery1 = rbinom(I, 1-exp(-gamma*dt));
  double recovery2 = rbinom(A, 1-exp(-gamma*dt));
  
  // Rprintf(\"trans_from_P[0] = %lg \\n\", trans_from_P[0]);
  // Rprintf(\"trans_from_P[1] = %lg \\n\", trans_from_P[1]);
  
  S += -infection;
  E += infection - E_to_P;
  P += E_to_P - trans_from_P[0] - trans_from_P[1];
  I += trans_from_P[0] - recovery1;
  A += trans_from_P[1]- recovery2;
  R += recovery1 + recovery2;
  Y += infection;
")

rmeas <- Csnippet("
  Z = Y;
")

dmeas <- Csnippet("
  lik = dpois(Z, Y+1e-6, give_log);
")

rinit <- Csnippet("
  S = 1e7;
  E = 0;
  P = 0;
  I = 10;  
  A = 0;
  R = 0;
  Y = 0;
  logbeta = log(0.24);
")

dat <- readRDS("outputs/Rt_dataset_20210831T201644.rds")
d <- data.frame(time = dat$time, Z = round(dat$ode$daily_infected))

sepiar <- pomp(data = d,
     times = "time",
     t0 = 0,
     rinit = rinit,
     rmeasure = rmeas,
     dmeasure = dmeas,
     rprocess = euler(sepiar_step, delta.t = 0.2),
     statenames = c("S","E","P","I","A","R","Y","logbeta"),
     paramnames = c("vol","gamma","ba","bp","fa","epsilon","delta"),
     accumvars = "Y")

```


```{r}
sepiar %>% simulate(params=c(gamma=0.4,vol=0.0002,ba=1,bp=1,
                             fa=0.3,epsilon=0.4,delta=0.2))%>% plot()
# Local search
set.seed(234)
sepiar %>%
  mif2(
    params=c(gamma=0.4,vol=0.2,ba=1,bp=1,fa=0.3,
             epsilon=0.4,delta=0.2),
    Np=10000,
    Nmif=100,
    partrans=parameter_trans(log=c("vol")),
    paramnames=c("vol"),
    rw.sd=rw.sd(vol=0.05),
    cooling.fraction.50=0.5) -> mf1

coef(mf1)
plot(mf1)

sepiar_sim <- sepiar %>% 
  simulate(nsim = 1e2, 
           params = c(vol=coef(mf1)[[2]],gamma=coef(mf1)[[1]],
                      ba=1,bp=1,fa=0.3,epsilon=0.4,delta=0.2)) 

## compare fits with the data
plot(sepiar_sim[[1]]@states["Y",], type="l", ylim=c(0,2000))
for(i in 2:100)  lines(sepiar_sim[[i]]@states["Y",])
points(sepiar@data["Z",], col=2, lwd=2)
```

```{r}
N <- sepiar_sim[[1]]@states["S",] + sepiar_sim[[1]]@states["E",] + 
  sepiar_sim[[1]]@states["P",] + sepiar_sim[[1]]@states["I",] + 
  sepiar_sim[[1]]@states["A",] + sepiar_sim[[1]]@states["R",] 


## beta
plot(exp(sepiar_sim[[1]]@states["logbeta",]), type="l", ylim=c(0,1))
for(i in 2:100)  lines(exp(sepiar_sim[[i]]@states["logbeta",]))

Rt <- data.frame(matrix(NA, nrow=201, ncol=100))
for(i in 1:100) {
  Rt[,i] <- 
    exp(sepiar_sim[[i]]@states["logbeta",])/0.2*sepiar_sim[[i]]@states["S",]/N
}
Rt$rowmean <- rowMeans(Rt)

Rt$t <- 0:200
plot(Rt$t, Rt$rowmean, type="l")
```




```{r}

pfsir <- pfilter(sir, Np=1000, paramnames = c("beta"), filter.mean = T, filter.traj = T)

plot(pfsir@data[1,])
lines(pfsir@filter.mean["I",], col=2)

## How can I beta time-varying and 
# created by the sir ode model with 
# library(Kmisc)
# y0 <- c(S = 1e7-10, I = 10, R = 0, CI = 10) # initial values
# params <- c(beta = 0.3, gamma = 0.2) # parameter values
# times <- seq(0, 50, by = 1) # daily output for 150 days
# library(deSolve)
# library(tidyverse)
# ode(y = y0, times = times, func = ode_sir, parms = params) %>% 
#   as.data.frame() -> out 
# daily_onset <- dput(round(c(0, diff(out$CI))))

daily_onset <- c(0, 3, 3, 4, 4, 5, 5, 6, 6, 7, 8, 9, 9, 10, 12, 13, 14, 16, 
17, 19, 21, 23, 26, 28, 31, 35, 38, 42, 47, 52, 57, 63, 70, 77, 
86, 95, 104, 115, 128, 141, 156, 172, 190, 210, 232, 257, 284, 
314, 346, 383, 423)

sir.step <- "
  double N = S + I + R;
  logbeta += rnorm(0, vol);
  double lambda = exp(logbeta) * I / N;
  double infection = rbinom(S,  1 - exp(-lambda * dt));
  double recovery = rbinom(I, 1 - exp(-gamma * dt));
  S += -infection;
  I += infection - recovery;
  R += recovery;
  Y += infection;
"

rmeas <-"
  Z = rpois(Y);
"

dmeas <-"
  lik = dpois(Z, Y, give_log);
"

rinit <-"
  S=1e7-10;
  I=10;
  R=0;
  Y=0;
  logbeta=log(0.3);
"

pomp(data = data.frame(time=seq(0,50,by=1),Z=daily_onset),
     times = "time",
     t0 = 0,
     rmeasure = Csnippet(rmeas),
     dmeasure = Csnippet(dmeas),
     rprocess = euler(Csnippet(sir.step), delta.t = 0.25),
     statenames = c("S","I","R","Y","logbeta"),
     paramnames = c("vol","gamma"), 
     rinit = Csnippet(rinit),
     accumvars="Y",
     params = c(vol=0.4,gamma=0.2)
) -> sir

pfsir <- pfilter(sir, Np=10000, paramnames = c("vol","gamma"), filter.mean = T, filter.traj = T)

plot(pfsir@data[1,])
lines(pfsir@filter.mean["Y",], col=2)
# beta
plot(exp(pfsir@filter.mean["logbeta",])*1e7/pfsir@filter.mean["S",])
```

### Iterated filtering
```{r}
library(pomp)
Csnippet("
  pop = rpois(b*N);  
  ") -> rmeas

Csnippet("
  N = N_0;
  ") -> rinit
Csnippet("
  double dW = rnorm(0,sqrt(dt));
  N += r*N*(1-N/K)*dt+sigma*N*dW;
") -> vpstepC

parus %>%
  pomp(
    times="year", t0=1960,
    rinit=rinit,
    rmeasure=rmeas,
    rprocess=euler(vpstepC,delta.t=1/365),
    statenames="N",
    paramnames=c("r","K","sigma","b","N_0")
  ) -> vpC

sobol_design(
  lower=c(r=0,K=100,sigma=0,N_0=150,b=1),
  upper=c(r=5,K=600,sigma=2,N_0=150,b=1),
  nseq=100
) -> guesses
plot(guesses,pch=16)

Csnippet("
  lik = dpois(pop,b*N,give_log);
") -> dmeas

vpC %>%
  mif2(
    params=z,
    Np=1000,
    Nmif=20,
    dmeasure=dmeas,
    partrans=parameter_trans(log=c("r","K","sigma","N_0")),
    rw.sd=rw.sd(r=0.02,K=0.02,sigma=0.02,N_0=ivp(0.02)),
    cooling.fraction.50=0.5,
    paramnames=c("r","K","sigma","N_0","b"),
    statenames=c("N")
  ) -> mf1
plot(mf1)

## estimating the log likelihood
replicate(5, mf1 %>% pomp::pfilter() %>% logLik()) %>% logmeanexp(se=TRUE)


library(parallel)
library(doParallel)
library(foreach)
ncores <- detectCores()

cl <- makeCluster(getOption("cl.cores", ncores - 2))
doParallel::registerDoParallel(cl)

foreach (guess=iter(guesses,"row"),
         .combine=c, .packages=c("pomp"),
         .errorhandling="remove", .inorder=FALSE) %dopar% {
        mf1 %>% mif2(params=guess)} -> mifs

parallel::stopCluster(cl)
library(tidyverse)
mifs %>%
  traces() %>%
  melt() %>%
  filter(variable!="b") %>%
  ggplot(aes(x=iteration,y=value,group=L1,color=L1))+
  geom_line()+
  facet_wrap(~variable,scales="free_y")+
  guides(color="none")


# Diagnostics in the last iteration of the filter
mifs %>% 
  as("data.frame") %>% 
  gather(variable,value,-year,-.id) %>%
  ggplot(aes(x=year,y=value,group=.id,color=.id))+
  geom_line()+
  facet_wrap(~variable,scales="free_y",ncol=1)+
  guides(color="none")

# Evaluate the likelihood of the estimates through several independent particle
# filter operation
cl <- makeCluster(getOption("cl.cores", ncores - 2))
doParallel::registerDoParallel(cl)

foreach (mf=mifs,
  .combine=rbind, .packages=c("pomp"), 
  .errorhandling="remove", .inorder=FALSE) %dopar% {
    
    replicate(5, 
      mf %>% pfilter() %>% logLik()
    ) %>%
      logmeanexp(se=TRUE) -> ll
    
    data.frame(as.list(coef(mf)),loglik=ll[1],loglik.se=ll[2])
    
  } -> estimates

parallel::stopCluster(cl)

# likelihood profile
estimates %>%
  filter(loglik>max(loglik)-10) %>%
  select(r,K,sigma,N_0,b) %>%
  apply(2,range) -> ranges
ranges

profile_design(
  r=10^seq(
    from=log10(ranges[1,1]),
    to=log10(ranges[2,1]),
    length=20
  ),
  lower=ranges[1,-1],
  upper=ranges[2,-1],
  nprof=50
) -> starts

dim(starts)

```


```{r}
library(pomp)
library(tidyverse)
seir_step <- Csnippet("
  double N = S + E1 + E2 + I + R;
  double dN_SE1 = beta * S * I / N * dt;
  double dN_E1E2 = 2 * sigma * E1 * dt;
  double dN_E2I = 2 * sigma * E2 * dt;
  double dN_IR = gamma * I * dt;
  S +=  - dN_SE1;
  E1 += dN_SE1 - dN_E1E2;
  E2 += dN_E1E2 - dN_E2I;
  I += dN_E2I - dN_IR;
  R += dN_IR;
  H += dN_SE1;
")

seir_rinit <- Csnippet("
  S = S_0;
  E1 = E1_0;
  E2 = E2_0;
  I = I_0;
  R = R_0;
  H = H_0;
")

seir_dmeas <- Csnippet("
  lik = dpois(daily_infect, H, give_log);
")

seir_rmeas <- Csnippet("
  daily_infect = rpois(H);
")

dat <- data.frame(day = Rt_data$t, daily_infect = Rt_data$daily_infect)

dat %>% pomp(
  times = "day", 
  t0 = 1,
  rprocess = euler(seir_step, delta.t = 0.2),
  rinit = seir_rinit,
  rmeasure = seir_rmeas,
  dmeasure = seir_dmeas,
  accumvars = "H",
  statenames = c("S", "E1", "E2", "I", "R", "H"),
  paramnames = c("beta", "sigma", "gamma", "S_0", "E1_0", "E2_0", "I_0", "R_0", "H_0")) -> covid_seir

covid_seir %>%
  simulate(params = c(beta=1.2*0.25, sigma=0.25, gamma=0.25,
                      S_0=1e7, E1_0=40, E2_0=40, I_0=100, R_0=0, H_0=0),
nsim = 20, format="data.frame", include.data=TRUE) -> sims

sims %>%
  ggplot(aes(x=day,y=daily_infect, group=.id, color=.id=="data"))+
  geom_line()+
  guides(color=FALSE)

covid_seir %>%
  pfilter(Np=1000, paramnames="beta") -> pfrick

```

```{r}
library(pomp)
library(tidyverse)
seir_step <- Csnippet("
  double N = S + E1 + E2 + I + R;
  double dN_SE1 = beta * S * I / N * dt;
  double dN_E1E2 = 2 * 0.25 * E1 * dt;
  double dN_E2I = 2 * 0.25 * E2 * dt;
  double dN_IR = 0.25 * I * dt;
  S +=  - dN_SE1;
  E1 += dN_SE1 - dN_E1E2;
  E2 += dN_E1E2 - dN_E2I;
  I += dN_E2I - dN_IR;
  R += dN_IR;
  H += dN_SE1;
")

seir_rinit <- Csnippet("
  S = 1e7;
  E1 = 40;
  E2 = 40;
  I = 100;
  R = 0;
  H = 0;
")

seir_dmeas <- Csnippet("
  lik = dpois(daily_infect, H, give_log);
")

seir_rmeas <- Csnippet("
  daily_infect = rpois(H);
")

dat <- data.frame(day = Rt_data$t, daily_infect = as.integer(Rt_data$daily_infect))

dat %>% pomp(
  times = "day", 
  t0 = 1,
  rprocess = euler(seir_step, delta.t = 0.2),
  rinit = seir_rinit,
  rmeasure = seir_rmeas,
  dmeasure = seir_dmeas,
  accumvars = "H",
  statenames = c("S", "E1", "E2", "I", "R", "H"),
  paramnames = c("beta")) -> covid_seir

covid_seir %>%
  simulate(params = c(beta=1.2*0.25),
nsim = 20, format="data.frame", include.data=TRUE) -> sims

sims %>%
  ggplot(aes(x=day,y=daily_infect, group=.id, color=.id=="data"))+
  geom_line()+
  guides(color=FALSE)

pf = pfilter(covid_seir, Np=1000, params = c(beta = 0.3), 
             filter.mean = T, filter.traj = T, verbose = T) 
pf@filter.mean %>% 
  t() %>%
  as.data.frame() %>%
  mutate(time = 1:n()) %>% 
  pivot_longer(-time) %>%
  filter(name == "I") %>% 
  ggplot(aes(time, value)) +
  geom_line(aes(color = name))
  
```      

```{r}
ewmeas %>%
  subset(time < 1952) %>%
  pomp(
    times="time",t0=1948,
    rprocess=euler(
      Csnippet("
        int nrate = 6;
        double rate[nrate];	// transition rates
        double trans[nrate];	// transition numbers
        double dW;

        // gamma noise, mean=dt, variance=(sigma^2 dt)
        dW = rgammawn(sigma,dt);

        // compute the transition rates
        rate[0] = mu*pop;	// birth into susceptible class
        rate[1] = (iota+Beta*I*dW/dt)/pop; // force of infection
        rate[2] = mu;		// death from susceptible class
        rate[3] = gamma;	// recovery
        rate[4] = mu;		// death from infectious class
        rate[5] = mu; 		// death from recovered class

        // compute the transition numbers
        trans[0] = rpois(rate[0]*dt);	// births are Poisson
        reulermultinom(2,S,&rate[1],dt,&trans[1]);
        reulermultinom(2,I,&rate[3],dt,&trans[3]);
        reulermultinom(1,R,&rate[5],dt,&trans[5]);

        // balance the equations
        S += trans[0]-trans[1]-trans[2];
        I += trans[1]-trans[3]-trans[4];
        R += trans[3]-trans[5];
      "),
      delta.t=1/52/20
    ),
    rinit=Csnippet("
      double m = pop/(S_0+I_0+R_0);
      S = nearbyint(m*S_0);
      I = nearbyint(m*I_0);
      R = nearbyint(m*R_0);
    "),
    paramnames=c("mu","pop","iota","gamma","Beta","sigma",
      "S_0","I_0","R_0"),
    statenames=c("S","I","R"),
    params=c(mu=1/50,iota=10,pop=50e6,gamma=26,Beta=400,sigma=0.1,
      S_0=0.07,I_0=0.001,R_0=0.93)
  ) -> ew1
```
